---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# plume

<!-- badges: start -->
[![R-CMD-check](https://github.com/arnaudgallou/plume/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/arnaudgallou/plume/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/arnaudgallou/plume/branch/main/graph/badge.svg)](https://app.codecov.io/gh/arnaudgallou/plume?branch=main)
<!-- badges: end -->

## Overview

plume provides tools for handling and generating author-related information for scientific writing in Rmarkdown and Quarto.
The package implements two `R6` classes:

- `Plume`: class that generates author lists and other author-related information
from tabular data.

- `PlumeQuarto`: class extending `Plume` that allows you to push or update author metadata in the yaml header of a Quarto file. The generated yaml complies with [Quarto's author and affiliations schemas](https://quarto.org/docs/journals/authors.html).

## Installation

You can install the development version of plume from GitHub with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("arnaudgallou/plume")
```

## Usage

The minimal required data to work with plume classes is a data set containing given and family names but you would normally want to provide more information such as email addresses, orcid, affiliations, etc.

```{r, message = FALSE}
library(plume)

encyclopedists

aut <- PlumeQuarto$new(encyclopedists)
aut
```

```{r, include = FALSE}
tmp <- withr::local_tempfile(
  lines = "---\ntitle: Encyclopédie\n---\n\nQui scribit bis legit",
  fileext = ".qmd"
)
```

`PlumeQuarto` lets you push author metadata in the yaml header of any `.qmd` files using `to_yaml()`.

Consider the following example:

```{r, echo = FALSE, comment = ""}
cat(readr::read_file(tmp))
```

```{r, eval = FALSE}
aut$set_corresponding_authors(1, 4)
aut$to_yaml("file.qmd")
```

```{r, echo = FALSE, comment = ""}
aut$set_corresponding_authors(1, 4)
aut$to_yaml(tmp)
cat(readr::read_file(tmp))
```

You can also generate author information as character strings:

```{r, message = FALSE}
aut <- Plume$new(encyclopedists, names = c(contribution = "contribution_n"))
aut$set_corresponding_authors(diderot, by = "family_name")

aut$get_author_list(format = "^a,^cn") |> enumerate(last = ",\n")

aut$get_contact_details()

aut$get_affiliations()

aut$get_notes()

aut$get_contributions()

aut_v <- Plume$new(
  encyclopedists,
  names = c(contribution = "contribution_v"),
  symbols = list(affiliation = letters)
)

aut_v$get_author_list(format = "^a^") |> enumerate(last = ",\n")

aut_v$get_contributions(role_first = FALSE, name_list = TRUE)
```

## Acknowledgements

Thanks to [Richard J. Telford](https://github.com/richardjtelford) for his advice
that helped me conceive this package.
