---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# plume

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/plume)](https://CRAN.R-project.org/package=plume)
[![R-CMD-check](https://github.com/arnaudgallou/plume/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/arnaudgallou/plume/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/arnaudgallou/plume/branch/main/graph/badge.svg)](https://app.codecov.io/gh/arnaudgallou/plume?branch=main)
<!-- badges: end -->

## Overview

plume provides tools for handling and generating author-related information for scientific writing in R Markdown and Quarto. The package implements two R6 classes:

- `PlumeQuarto`: class that allows you to push author metadata in the YAML header of Quarto files. The generated YAML complies with Quarto's [author and affiliations schemas](https://quarto.org/docs/journals/authors.html). This is the class to use if you work with journal templates.

- `Plume`: class that generates author lists and other author-related information as character strings. This is an easy and convenient solution when you don't need preformatted documents.

## Installation

Install plume from CRAN with:

```{r, eval = FALSE}
install.packages("plume")
```

Alternatively, you can install the development version of plume from GitHub with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("arnaudgallou/plume")
```

## Usage

The minimal required data to work with plume classes is a data set containing given and family names but you would normally want to provide more information such as email addresses, ORCIDs, affiliations, etc.

```{r, message = FALSE}
library(plume)

encyclopedists

Plume$new(encyclopedists)
```

```{r, include = FALSE}
tmp_file <- withr::local_tempfile(
  lines = "---\ntitle: Encyclopédie\n---\n\nQui scribit bis legit",
  fileext = ".qmd"
)
```

`PlumeQuarto` lets you push author metadata in the YAML header of any `.qmd` file using the `to_yaml()` method.

Consider the following example:

```{r, echo = FALSE, comment = ""}
cat(readr::read_file(tmp_file))
```

```{r, eval = FALSE}
aut <- PlumeQuarto$new(
  encyclopedists,
  file = "file.qmd"
)
aut$set_corresponding_authors(1, 4)
aut$to_yaml()
```

```{r, echo = FALSE, comment = ""}
aut <- PlumeQuarto$new(encyclopedists, tmp_file)
aut$set_corresponding_authors(1, 4)
aut$to_yaml()
cat(readr::read_file(tmp_file))
```

Alternatively, you can generate author information as character strings using `Plume`:

```{r, message = FALSE}
aut <- Plume$new(encyclopedists)
aut$set_corresponding_authors(diderot, .by = "family_name")

aut$get_author_list(suffix = "^a,^cn")

aut$get_contact_details()

aut$get_affiliations()

aut$get_notes()

aut$get_contributions()

aut2 <- Plume$new(
  encyclopedists,
  roles = c(
    supervision = "supervised the project",
    writing = "contributed to the Encyclopédie"
  ),
  symbols = list(affiliation = letters)
)

aut2$get_author_list("^a^")

aut2$get_contributions(roles_first = FALSE, divider = " ")
```

## Acknowledgements

Thanks to:

* [Richard J. Telford](https://github.com/richardjtelford) for his advice that helped me conceive this package.

* [Maëlle Salmon](https://github.com/maelle) and [Gábor Csárdi](https://github.com/gaborcsardi) for their help when I was stuck with unit tests, roxygen2 or pkgdown.
