---
title: "plume workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plume workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

plume is designed to work with spreadsheets, which makes it easy to maintain and share author data with co-authors.

I recommend using [Google Sheets](https://docs.google.com/spreadsheets/) for this. The [googlesheets4](https://googlesheets4.tidyverse.org) package can help us manage everything from R.

First, we need to create a spreadsheet. We can use `plume_template()` to use the default plume spreadsheet template.

Note that if it's the first time you use googlesheets4, you'll need to grant the package permission to operate with Google Sheets. You can read more about googlesheets4 authentication [here](https://googlesheets4.tidyverse.org/reference/gs4_auth.html).

```{r eval = FALSE}
library(googlesheets4)

gs4_create(
  name = "encyclopédie",
  sheets = plume_template()
)
```

The above will create a spreadsheet named `encyclopédie` with the columns defined by `plume_template()`.

Enter your information and share the spreadsheet to your collaborators. You can do it either from Google Drive or in R using `googledrive::drive_share()`.

Once the spreadsheet contains all author metadata, we can use `googlesheets4::read_sheet()` to read in the spreadsheet in R.

```{r eval = FALSE}
read_sheet(gs4_find("encyclopédie"))
```

Below is the typical structure we would use in a manuscript written in R Markdown or Quarto.

Note that this code would have to be put in a separate R script when using `PlumeQuarto` to inject author metadata.

````{verbatim}
---
title: An awesome future publication
---
  
```{r}
#| label: setup
#| include: false

library(googlesheets4)
library(plume)

tbl_authors <- gs4_find("plume authors") |> read_sheet()

aut <- Plume$new(tbl_authors)
aut$set_corresponding_authors(1)
```
````

To modify author data, simply edit the spreadsheet. All author information will be automatically updated the next time you render the document.
