---
title: "plume workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plume workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

plume is designed to work with spreadsheets, which makes it easy to store, maintain and share data with co-authors. I recommend using [Google Sheets](https://docs.google.com/spreadsheets/) because it works great with the [googlesheets4](https://googlesheets4.tidyverse.org) package to manage sheets from R.

First, you need to create a spreadsheet. `plume_template()` provides a default spreadsheet template for that purpose.

Note that if it's the first time you use googlesheets4, you'll need to grant the package permission to operate with Google Sheets. Read more about googlesheets4 authentication [here](https://googlesheets4.tidyverse.org/reference/gs4_auth.html).

```{r eval = FALSE}
library(googlesheets4)

gs4_create(
  name = "encyclopédie",
  sheets = plume_template()
)
```

The above will create a sheet named `encyclopédie` with the columns defined by `plume_template()`.

Enter your information and share the sheet to your collaborators so they can add theirs too. Once the sheet contains all author metadata, use `read_sheet()` to read it in R.

```{r eval = FALSE}
read_sheet(gs4_find("encyclopédie"))
```

Below is what the typical plume set up would look like in a R markdown or Quarto document.

Note that the code would have to be put in a separate R script when using `PlumeQuarto` to inject author metadata.

````{verbatim}
---
title: An awesome future publication
---
  
```{r}
#| label: setup
#| include: false

library(googlesheets4)
library(plume)

tbl_authors <- gs4_find("encyclopédie") |> read_sheet()

aut <- Plume$new(tbl_authors)
aut$set_corresponding_authors(1)
```
````

To modify author data, simply edit the spreadsheet. All author information in the document will update automatically the next time you render it.
