---
title: "Using CRediT roles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using CRediT roles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(plume)
```

plume implements a simple solution to work with the `r plume:::link("crt")` (CRediT).

First, create a table using `plm_template(credit_roles = TRUE)` to ensure the table has all the appropriate columns. CRediT columns are standardised and can't be changed so it's important that they're properly named.

Use a `1` to assign a role to a specific contributor and leave it blank otherwise. All contributions should normally be used. For the purpose of the example, I selected just a few columns here.

```{r}
#| echo: false
tbl_credit <- encyclopedists |> 
  dplyr::select(given_name, family_name) |> 
  dplyr::mutate(
    conceptualization = c(1, NA, NA, NA),
    analysis = rep(1, 4),
    supervision = c(1, NA, NA, 1),
    methodology = rep(1, 4),
    writing = rep(1, 4)
  )

tbl_credit
```

Once the data set is ready, simply use `credit_roles = TRUE` when creating your `plume` object. plume will process and format the roles automatically.

```{r}
aut <- Plume$new(tbl_credit, credit_roles = TRUE)
aut$get_contributions()
```

